An Open Repository of Real-Time COVID-19 Indicators

### 2021-09-16 indicators/paper/responses-to-reviews.tex

\begin{quote}
  I think that descriptions in results of ``what the data could be used for''
  are less compelling than more specific examples would be. For instance, the
  section on revised estimates over time seems to echo what was in the methods
  and implications don't really hit home without connection to changes in
  specific forecasts.
\end{quote}

TODO

\begin{quote}
  Some of the verb tenses (e.g. line 33) feel awkward: ``The Delphi Group
  **works** (worked) with partner organizations and34 public data sets to build
  a massive database of indicators35 tracking COVID-19 activity and other
  relevant phenomena36 in the United States, which has been publicly available
  and37 continuously updated since April 2020.''
\end{quote}

We have tweaked this sentence and a few others, though we note that the present
tense is accurate in many cases, because data collection and aggregation is
ongoing.

@@

\begin{quote}
  Similarly, the verb ``ingest'' also feels odd to refer to data sources.
\end{quote}

TODO

\begin{quote}
  I'm not sure that it is common practice to repeatedly refer to a sponsoring
  research group throughout an academic article.
\end{quote}

We have adjusted the language used throughout, so references to the research
group are limited to where they are most useful to the reader.

\begin{quote}
  Descriptions of statistical analyses (e.g. correlation methods) feel more
  appropriate for the methods section, not the results section.
\end{quote}
We are open to changes if the Editor feels they're necessary to match typical
\textit{PNAS} style, but we feel that in our paper, it makes the most sense for
the Methods to describe the methods underlying the data we gather and aggregate,
since that data is the main focus of the article. The Results focus on
illustrations of the data's utility, including their correlations with other
COVID indicators.

### 2021-09-16 indicators/paper/paper.tex
Language tweaks requested by Reviewer 2

The Delphi Group works with partner organizations and public data sets to build
a massive database of indicators tracking COVID-19 activity and other relevant
phenomena in the United States, which has been publicly available and
continuously updated since April 2020. Alongside public data on reported cases
and deaths, this database includes several unique data streams, including
indicators extracted from de-identified medical claims data, antigen test
results from a major testing manufacturer, large-scale public surveys that
measure symptoms and public behavior, and indicators based on particular Google
search queries. (We use the terms ``indicator'' and ``signal'' interchangeably.)
We make aggregate signals publicly available, generally at the county level, via
the COVIDcast API \cite{CovidcastAPI}. We store and provide access to all
previous (historical) versions of the signals, a key feature that exposes the
effects of data revisions.  Lastly, we provide R \cite{CovidcastR} and Python
\cite{CovidcastPy} packages to facilitate interaction with the API, and an
online dashboard to visualize the data \cite{CovidcastViz}.

@@

\begin{table*}[t]
\centering
\caption{Data sources available in Delphi's COVIDcast API \cite{CovidcastAPI},
  as of date of publication. The first group of data sources are produced by
  Delphi from data not otherwise available publicly (or only available in
  limited form); the second group is mirrored from public sources.}
\begin{tabular}{>{\raggedright}p{1.2in} p{4.0in} l >{\raggedright\arraybackslash}p{0.5in}}
  \toprule
  \textbf{Data source} & \textbf{Signals available} & \textbf{First date} &
\textbf{Resolution} \\\midrule
  % Exclusive data
  Change Healthcare & Percentage of outpatient visits with COVID diagnostic
codes or codes indicating COVID-like symptoms. Based on de-identified claims
data processed by Change Healthcare. & 2020-02-01 & County$^*$ \\
  Doctor Visits & Percentage of outpatient visits primarily about COVID-like
symptoms, based on de-identified claims data provided by health system
partners. & 2020-02-01 & County$^*$ \\
  Hospital Admissions & Percentage of new hospital admissions with COVID
diagnostic codes, based on de-identified claims data provided by health system
partners. & 2020-02-01 & County$^{**}$\\
  Quidel & Test positivity rates for COVID-19 antigen tests produced by
Quidel. & 2020-05-26 & County$^{**}$ \\
  SafeGraph & Mobility metrics, such as time away from home or visits to bars
and restaurants, based on cell phone mobility data collected by SafeGraph
\cite{SafeGraphSocial, SafeGraphPatterns}. & 2019-01-01 & County \\
  COVID-19 Trends and Impact Survey & COVID symptoms, social distancing
behaviors, mental health, economic impact, behavior (e.g., mask-wearing,
vaccination attitudes), and COVID testing signals based on daily surveys
conducted nationally by Delphi through Facebook \cite{DelphiSurvey,
Kreuter:2020}. & 2020-04-06 & County$^{**}$ \\
  \midrule
  % Mirrored data from other sources
  Health \& Human Services & Counts of hospital admissions due to confirmed or
suspected COVID-19, as reported by the Department of Health \& Human Services. &
2019-12-31 & State \\
  COVID Act Now & COVID-19 testing results, such as positivity rate and number
of tests, compiled by COVID Act Now from CDC reporting. & 2020-03-02 &
County$^*$ \\
  Google Symptoms & Trends in Google search volume for terms related to
anosmia and ageusia (loss of smell or taste), which correlate with COVID
activity, based on data shared by Google \cite{GoogleSymptoms}. & 2020-02-13 &
County$^{***}$ \\
  Cases and Deaths & Confirmed COVID-19 cases and deaths, compiled by JHU CSSE
\cite{Dong:2020} and by USAFacts \cite{USAFacts}. & 2020-01-22 & County \\
    NCHS Mortality & Weekly totals of deaths broken down by cause, such as
COVID, flu, or pneumonia, compiled by the National Center for Health Statistics
\cite{NCHS}. & 2020-01-26 & State \\
  \bottomrule
\end{tabular}
\addtabletext{$^*$Available at $>60$\% of counties; $^{**}$available at
  20--60\% of counties; $^{***}$available at $<20$\% of counties.  For some
  signals, location availability varies over time, e.g., due to variable
  reporting volume.}
\label{tab:sources_signals}
\end{table*}

@@

For this reason, the COVIDcast API annotates every observation with two dates:
the \textit{time value}, the date the underlying events (such as tests or
doctor's visits) occurred, and the \textit{issue date} when Delphi aggregated
and reported the data for that time value. Importantly, there can be multiple
observations for a single time value with different issue dates, for example if
data is revised or claims records arrive late. Delphi tracks revisions to all
data sources we ingest, including external data sources (such as sources
tracking cases and deaths). Many external sources do not keep a public or
conveniently accessible record of revisions of their data.

@@

Since April 2020, Delphi has been maintaining and continually improving various
online visualization tools for the COVIDcast indicators \cite{CovidcastViz}.
These tools fetch data directly from the API, and allow for exploration of both
temporal (e.g., time series graphs) and spatial (e.g., choropleth maps) trends
in the signals, as well as many other aspects, such as correlations, anomalies,
and backfill. There is also a dedicated dashboard for exploring results from the
COVID-19 Trends and Impact Survey. The visualizations have been continually
improved as new sources of data arrive, and in response to interviews with users
and health experts, usage analytics from the site, and user surveys. Key design
goals for visualizations displaying COVIDcast data have included representing
uncertainty and missing data, making features consistent and discoverable
despite a wide variety of signal types, and smoothly switching between different
geographic aggregations and time scales.

@@

Since July 2020, Delphi has been regularly submitting short-term forecasts of
COVID-19 case and death incidence, at the state and county levels, to the
COVID-19 Forecast Hub \cite{ForecastHub}, with ``CMU-TimeSeries'' as the
team-model name.  The process of building, training, and deploying our
forecasting models leverages much of the infrastructure described in this paper
(such as the COVIDcast API's \textit{as of} feature), and some of our
forecasting systems rely on auxiliary indicators (such as survey-based and
claims-based COVID-like illness signals, which are described below).

>>>>

The Delphi Group worked with partner organizations and public data sets to build
a massive database of indicators tracking COVID-19 activity and other relevant
phenomena in the United States, which has been publicly available and
continuously updated since April 2020. Alongside public data on reported cases
and deaths, this database includes several unique data streams, including
indicators extracted from de-identified medical claims data, antigen test
results from a major testing manufacturer, large-scale public surveys that
measure symptoms and public behavior, and indicators based on particular Google
search queries. (We use the terms ``indicator'' and ``signal'' interchangeably.)
We make aggregate signals publicly available, generally at the county level, via
the COVIDcast API \cite{CovidcastAPI}. We store and provide access to all
previous (historical) versions of the signals, a key feature that exposes the
effects of data revisions.  Lastly, we provide R \cite{CovidcastR} and Python
\cite{CovidcastPy} packages to facilitate interaction with the API, and an
online dashboard to visualize the data \cite{CovidcastViz}.

@@

\begin{table*}[t]
\centering
\caption{Data sources available in the COVIDcast API \cite{CovidcastAPI}, as of
  date of publication. The first group of data sources are produced from data
  not otherwise available publicly (or only available in limited form); the
  second group is mirrored from public sources.}
\begin{tabular}{>{\raggedright}p{1.2in} p{4.0in} l >{\raggedright\arraybackslash}p{0.5in}}
  \toprule
  \textbf{Data source} & \textbf{Signals available} & \textbf{First date} &
\textbf{Resolution} \\\midrule
  % Exclusive data
  Change Healthcare & Percentage of outpatient visits with COVID diagnostic
codes or codes indicating COVID-like symptoms. Based on de-identified claims
data processed by Change Healthcare. & 2020-02-01 & County$^*$ \\
  Doctor Visits & Percentage of outpatient visits primarily about COVID-like
symptoms, based on de-identified claims data provided by health system
partners. & 2020-02-01 & County$^*$ \\
  Hospital Admissions & Percentage of new hospital admissions with COVID
diagnostic codes, based on de-identified claims data provided by health system
partners. & 2020-02-01 & County$^{**}$\\
  Quidel & Test positivity rates for COVID-19 antigen tests produced by
Quidel. & 2020-05-26 & County$^{**}$ \\
  SafeGraph & Mobility metrics, such as time away from home or visits to bars
and restaurants, based on cell phone mobility data collected by SafeGraph
\cite{SafeGraphSocial, SafeGraphPatterns}. & 2019-01-01 & County \\
  COVID-19 Trends and Impact Survey & COVID symptoms, social distancing
behaviors, mental health, economic impact, behavior (e.g., mask-wearing,
vaccination attitudes), and COVID testing signals based on daily surveys
conducted nationally by Delphi through Facebook \cite{DelphiSurvey,
Kreuter:2020}. & 2020-04-06 & County$^{**}$ \\
  \midrule
  % Mirrored data from other sources
  Health \& Human Services & Counts of hospital admissions due to confirmed or
suspected COVID-19, as reported by the Department of Health \& Human Services. &
2019-12-31 & State \\
  COVID Act Now & COVID-19 testing results, such as positivity rate and number
of tests, compiled by COVID Act Now from CDC reporting. & 2020-03-02 &
County$^*$ \\
  Google Symptoms & Trends in Google search volume for terms related to
anosmia and ageusia (loss of smell or taste), which correlate with COVID
activity, based on data shared by Google \cite{GoogleSymptoms}. & 2020-02-13 &
County$^{***}$ \\
  Cases and Deaths & Confirmed COVID-19 cases and deaths, compiled by JHU CSSE
\cite{Dong:2020} and by USAFacts \cite{USAFacts}. & 2020-01-22 & County \\
    NCHS Mortality & Weekly totals of deaths broken down by cause, such as
COVID, flu, or pneumonia, compiled by the National Center for Health Statistics
\cite{NCHS}. & 2020-01-26 & State \\
  \bottomrule
\end{tabular}
\addtabletext{$^*$Available at $>60$\% of counties; $^{**}$available at
  20--60\% of counties; $^{***}$available at $<20$\% of counties.  For some
  signals, location availability varies over time, e.g., due to variable
  reporting volume.}
\label{tab:sources_signals}
\end{table*}

@@

For this reason, the COVIDcast API annotates every observation with two dates:
the \textit{time value}, the date the underlying events (such as tests or
doctor's visits) occurred, and the \textit{issue date} when we aggregated
and reported the data for that time value. Importantly, there can be multiple
observations for a single time value with different issue dates, for example if
data is revised or claims records arrive late. We track revisions to all
data sources we ingest, including external data sources (such as sources
tracking cases and deaths). Many external sources do not keep a public or
conveniently accessible record of revisions of their data.

@@

Since April 2020, we have been maintaining and continually improving various
online visualization tools for the COVIDcast indicators \cite{CovidcastViz}.
These tools fetch data directly from the API, and allow for exploration of both
temporal (e.g., time series graphs) and spatial (e.g., choropleth maps) trends
in the signals, as well as many other aspects, such as correlations, anomalies,
and backfill. There is also a dedicated dashboard for exploring results from the
COVID-19 Trends and Impact Survey. The visualizations have been continually
improved as new sources of data arrive, and in response to interviews with users
and health experts, usage analytics from the site, and user surveys. Key design
goals for visualizations displaying COVIDcast data have included representing
uncertainty and missing data, making features consistent and discoverable
despite a wide variety of signal types, and smoothly switching between different
geographic aggregations and time scales.

@@

Since July 2020, we have been regularly submitting short-term forecasts of
COVID-19 case and death incidence, at the state and county levels, to the
COVID-19 Forecast Hub \cite{ForecastHub}, with ``CMU-TimeSeries'' as the
team-model name.  The process of building, training, and deploying our
forecasting models leverages much of the infrastructure described in this paper
(such as the COVIDcast API's \textit{as of} feature), and some of our
forecasting systems rely on auxiliary indicators (such as survey-based and
claims-based COVID-like illness signals, which are described below).
